Novelty Detection Using Graphical Models for Semantic Room Classification

### 2011-10-10 semantic-nd/presentation.tex
Presentation for EPIA-2011 (finished maybe).

\documentclass[compress]{beamer}
\usepackage[utf8x]{inputenc}
\usepackage{default}
\usepackage{graphics}
\usepackage{ulem}
\usepackage{multicol}

@@

\section{Summary}
\begin{frame}{Summary - Plan}
    \begin{block}{Semantic mapping of indoor enviroments}
        \begin{itemize}
            \item What is it? What for?
            \item How to, using graphical models
            \item Limited on the agent knowledge of variable categories
        \end{itemize}
    \end{block}
    \begin{block}{Novelty Detection}
        \begin{itemize}
            \item Trigger on $P(x|known)$
            \item Dynamic graphs
            \item $P(x|known)/P(x)$
            \item Approximating $P(x)$
        \end{itemize}
    \end{block}
\end{frame}

\begin{frame}{Goals}
  Brief introduction
  \begin{itemize}
    \item Study a semantic mapping process in the context of mobile robots.
    \item Propose a method based on the studied system to detect novel semantic categories of places.
  \end{itemize}
\end{frame}

\begin{frame}{Outline}
  \begin{multicols}{2}
    \tableofcontents
  \end{multicols}
\end{frame}

@@

\subsection{Dora Architecture Overview}
\begin{frame}{Dora Architecture Overview}
    \begin{block}{Multi-modal approach}
        Combining semantic data extracted from diferent sources is expected to yield better performance.
    \end{block}
    \begin{block}{Probabilistic approach}
        TODO: Need to defend use of probabilistic approach and graphical models.
    \end{block}
\end{frame}

@@

\subsection{Spatial Knowledge and Conceptual Map}
\begin{frame}{Spatial Knowledge and Conceptual Map}
  \begin{columns}[c]
  \column{0.5\textwidth}
    \begin{figure}
      \includegraphics[width=\textwidth]{figures/chaingraph.pdf}
    \end{figure}
  \column{0.5\textwidth}
    \begin{block}{Conceptual Map}
      The ontology is used to identify instances of the concepts and categories
      and perform probabilistic semantic mapping using graphical models.
    \end{block}
  \end{columns}
\end{frame}

\subsection{Graphical Models}
\begin{frame}{Graphical models}
    Multi-modal probabilistic semantic mapping using graphical models.
    Uncertain semantic information
    Conceptual map
    Probabilistic inference on variables

            $\text{known} = \{\text{corridor}, \text{office}, \cdots\}$
            $R_i \in \text{known}$

    TODO: picture explaining that maybe roomA is not known and showing X (observed features) and introduced novel and not novel, and showing graph structure.

\end{frame}

@@

\subsection{Novelty Detection via Thresholding}
\begin{frame}
    $\frac{P(x|\overline{novel})P(\overline{novel})}{P(x)} \propto \frac{P(x|\overline{novel})}{P(x)} \propto P(x|\overline{novel})$
    \begin{block}{$P(novel)$ constant}
        Warning: assuming any graph structure has the same probability of generating
        novel classes.
    \end{block}
    \begin{block}
        Assumption on uniform $P(x)$.
    \end{block}
\end{frame}

\subsection{Dynamic distributions}
\begin{frame}
    The set of observed features and graph structure is always changing.
    Threshold on $P(x|known)$ \cite{bishop1994novelty}
    Dynamic graphs
\end{frame}

\subsection{Semi-supervised}
\begin{frame}

\end{frame}

\subsection{Results}
\begin{frame}
    Synthetic dataset where 1 room category is responsible for generating a variable set of observed features.
\end{frame}

\section{Conclusion}
\subsection{Conclusion}
\begin{frame}
    \begin{block}
        {Semantic Mapping Process}
        A multi-modal semantic mapping process based on \emph{probabilistic graphical models}.
    \end{block}
    \begin{block}
        {Novelty detection via thresholding}
        Presented a method to detect novel classes on a variable of \emph{probabilistic graphical models}
        dinamicaly instatiated from sensed semantic data.
    \end{block}
\end{frame}

\subsection{Future Work}
\begin{frame}{Future Work}
    \begin{block}
        {Incorporate novelty signals}
    \end{block}
\end{frame}

>>>>

\documentclass[compress]{beamer}
\usepackage[utf8x]{inputenc}
\usepackage{default}
\usepackage{graphics}
\usepackage{ulem}
\usepackage{multicol}
\usepackage{subfig}

@@

\section{Summary}
%\begin{frame}{Summary - Plan}
%    \begin{block}{Semantic mapping of indoor enviroments}
%        \begin{itemize}
%            \item What is it? What for?
%            \item How to, using graphical models
%            \item Limited on the agent knowledge of variable categories
%        \end{itemize}
%    \end{block}
%    \begin{block}{Novelty Detection}
%        \begin{itemize}
%            \item Trigger on $P(x|known)$
%            \item Dynamic graphs
%            \item $P(x|known)/P(x)$
%            \item Approximating $P(x)$
%        \end{itemize}
%    \end{block}
%\end{frame}

\subsection{Context}
\begin{frame}{Context}
  Brief introduction
  \begin{itemize}
    \item Study a semantic mapping process in the context of mobile robots.
    \item Propose a method based on the studied system to detect novel semantic categories of places.
  \end{itemize}
\end{frame}

\subsection{Outline}
\begin{frame}{Outline}
  \begin{multicols}{2}
    \tableofcontents
  \end{multicols}
\end{frame}

@@

\subsection{Dora Architecture Overview}
\begin{frame}{Dora Architecture Overview}
    \begin{block}{Multi-modal approach}
        Combining semantic data extracted from diferent sources is expected to yield better performance.
    \end{block}
    \begin{block}{Probabilistic approach}
        Uses probabilistic approach in classification in order to handle uncertainty.
    \end{block}
\end{frame}

@@

\subsection{Spatial Knowledge and Conceptual Map}
\begin{frame}{Spatial Knowledge and Conceptual Map}
    Using the spatial knowledge the agent instatiates a conceptual map, here represented as a chain-graph:
    \begin{figure}
      \includegraphics[width=0.5\textwidth]{figures/chaingraph.pdf}
    \end{figure}
\end{frame}

@@

\subsection{Novelty Detection via Thresholding}
\begin{frame}{Thresholding}
    \begin{block}
        {Optimal detector}
        \begin{itemize}
            \item Threshold on $P(R_i \in \text{known}|x) = P(\overline{novel}|x)$
            \item Optimizes true-positives given a fixed false-positive ratio.
        \end{itemize}
    \end{block}
    \begin{block}
        {Usually aproximated by $P(x|\overline{novel})$}
        \begin{itemize}
            \item Bayes rule
            \item Assumption on constant $P(\overline{novel})$
            \item Assumption on constant $P(x)$
        \end{itemize}

        \begin{equation*}
            \frac{P(x|\overline{novel})P(\overline{novel})}{P(x)} \propto \frac{P(x|\overline{novel})}{P(x)} \propto P(x|\overline{novel})
        \end{equation*}
    \end{block}
\end{frame}

\begin{frame}{Dynamic distributions}
    \begin{columns}[c]
        \column{0.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/conditional-prob-graph.pdf}

        \column{0.6\textwidth}
        \begin{itemize}
            \item Distribution changes with graph structure and sensed features.
            \item A threshold trained on $P(x|\overline{novel})$ is only valid for a specific
                graph.
        \end{itemize}
    \end{columns}
    \begin{center}
        \uncover<2>{\alert{$P(x)$ is not constant - needs to be approximated.}}
    \end{center}
\end{frame}

\subsection{Unconditional Probability}
\begin{frame}{Unconditional Probability}
    Assume $P(x)$ to be uniform:
    \vfill
    \begin{figure}
        \centering
        \includegraphics[width=0.50\textwidth]{figures/uniform-prob-graph.pdf}
    \end{figure}
    \vfill
    Expected to approximate poorly as it discards knowledge on the graph structure and other variables.
\end{frame}

\begin{frame}{Unconditional Probability}
    \only<1>{
    Unability to model $R_1$ leads to all factors dependent on it become between themselves:
    \vfill
    \begin{figure}
        \centering
        \includegraphics[width=0.50\textwidth]{figures/unconditional-prob-graph.pdf}
    \end{figure}
    }
    \only<2>{
    Approach the unknown factor with independent single factors that can easily be trained from unlabelled data:
    \vfill
    \begin{figure}
        \centering
        \includegraphics[width=0.50\textwidth]{figures/single-unconditional-prob-graph.pdf}
    \end{figure}
    }
\end{frame}

\subsection{Synthetic Distribution}
\begin{frame}
    {Results}
    \begin{itemize}
        \item Synthetic distribution.
        \item 11 diferent room categories and 9 types of room properties.
        \item Database of properties build from web data.
        \item A variable number of independent features is generated from a given room category.
        \item Only 5 room categories are given on labelled data.
    \end{itemize}
    \only<1>{
        \begin{figure}[h]
            \centering
            \includegraphics[width=0.40\textwidth]{figures/simple-cond-graph.pdf}
        \end{figure}
    }
    \only<2>{
        \begin{figure}[h]
            \centering
            \subfloat[Uniform approach]{\includegraphics[width=0.30\textwidth]{figures/simple-uniform-graph.pdf}}
            \qquad \qquad
            \subfloat[Independent approach]{\includegraphics[width=0.30\textwidth]{figures/simple-independent-graph.pdf}}
        \end{figure}
    }
\end{frame}

\subsection{Results}
\begin{frame}{Results}
    \begin{figure}[h]
    \centering
    \includegraphics[width=0.60\textwidth]{results/synthetic-all.pdf}

    \caption{\label{fig:synthetic-roc}ROC curve comparing novelty detection performance
             under samples with variable size of sensed properties.}
    \end{figure}
\end{frame}

\begin{frame}{Results}
    \begin{figure}[h]
        \centering

        \subfloat[3 sensed features]{\includegraphics[width=0.35\textwidth]{results/synthetic-3features.pdf}}
        \qquad
        \subfloat[5 sensed features]{\includegraphics[width=0.35\textwidth]{results/synthetic-5features.pdf}}

        \subfloat[10 sensed features]{\includegraphics[width=0.35\textwidth]{results/synthetic-10features.pdf}}
        \qquad
        \subfloat[50 sensed features]{\includegraphics[width=0.35\textwidth]{results/synthetic-50features.pdf}}

%        \caption{\label{fig:synthetic-roc-breakdown}ROC curves plotted showing performance of the
%        presented novelty detection method for graphs generated for different amount of
%        sensed features.}
    \end{figure}
\end{frame}

\section{Conclusion}
\subsection{Summary}
\begin{frame}{Summary}
    \begin{itemize}
        \item Studies a semantic mapping process
        \item Studies novelty detection
        \item Proposes a method to perform detection of novel categories
              of a variable when the graph structure is dynamic.
        \item Shows access to unlabelled data can greatly increase performance.
    \end{itemize}
\end{frame}

\subsection{Limitations}
\begin{frame}{Limitations and Future Work}
    \begin{block}{Assumption on constant $P(novel)$}
        Graph structure plays a role on the likelihood a sample with a novel
        category is drawn. Assuming it to be constant is a very strong assumption
        that is not expected to hold in all scenarios.
    \end{block}
    \begin{block}
        {Generalize the framework}
        Handle probabilistic classification from low-level classifiers and incorporate novelty signals from them.
    \end{block}
\end{frame}

\subsection{Extra Information}
\begin{frame}{Extra}
    \begin{block}{Reproducible Research}
        All reports and results reproducible from online repository:
        \url{https://github.com/andresusanopinto/novelty-detection-thesis}
    \end{block}
\end{frame}
